Combining information from photometric and spectroscopic data:\\ Don't waste your time in spectrophotometric calibration!

### 2014-04-28 tex/ms.tex
cleaning up mean model definition

If you don't believe the reported data noise variances $\sigma_n^2$,
it often makes sense to add in quadrature a ``jitter'' variance $s^2$.
This is a trivial modification of the chi-squared log-likelihood:
\begin{eqnarray}\label{eq:photometricLF}
\ln p(d\given\theta,s^2) &=& -\frac{1}{2}\,\sum_{n=1}^N \left[\frac{[d_n - \mu_n(\theta)]^2}{[\sigma_n^2 + s^2]} + \ln(2\pi\,[\sigma_n^2 + s^2]) \right]
\quad .
\end{eqnarray}
This modification shows the necessity of including the log-variance
term.
This---equation~(\ref{eq:photometricLF})---is in fact the log-likelihood
function we are going to use for the photometric measurements (BDJ: IS
THIS TRUE?) in what follows.
In this formulation, if $\theta$ are the parameters of interest, the
jitter variance $s^2$ is a nuisance noise-model parameter.
The Gaussian Process will be a generalization of this modification.

@@

We are now left to demonstrate that it is possible to infer the
parameters of interest $\theta$ while marginalizing over the
calibration parameters $\alpha$.

As a demonstration and test of the methodology for spectrophotometric
calibration described above, we will apply it to an \emph{actual,
  real} optical spectrum of a stellar cluster in M31.  Spectra of star
clusters provide a good test case, as they are less complex than
galaxy spectra (for which it is necessary to model the star formation
history) but are by no means trivial. In this way the key features of the
calibration model and methodology are not obscured, but the influence
of the .

We model the stellar cluster as a single age SSP. The physical
parameters $\theta$ which affect the spectrum of a star
cluster include the age, metallicity, stellar mass, attenuation by dust,
velocity dispersion, redshift, and uncertain
aspects of the stellar population synthesis models, including the
intial mass function.

The model for the spectrum is then 
\begin{eqnarray}
F(\lambda_o, \theta) = G(\sigma_v) \ast m_* S(\lambda_e/(1+z), Z, t,
imf [, {\rm etc}]) e^ {-k_\lambda \tau_V}
\end{eqnarray}

BDJ INSERT WORDS HERE.

@@

BDJ INSERT WORDS HERE.

>>>>

If you don't believe the reported data noise variances $\sigma_n^2$,
it often makes sense to add in quadrature a ``jitter'' variance $s^2$.
This is a trivial modification of the chi-squared log-likelihood:
\begin{eqnarray}\label{eq:photometricLF}
\ln p(d\given\theta,s^2) &=& -\frac{1}{2}\,\sum_{n=1}^N \left[\frac{[d_n - \mu_n(\theta)]^2}{[\sigma_n^2 + s^2]} + \ln(2\pi\,[\sigma_n^2 + s^2]) \right]
\quad .
\end{eqnarray}
This modification shows the necessity of including the log-variance
term.
This---equation~(\ref{eq:photometricLF})---is in fact the log-likelihood
function we are going to use for the photometric measurements (BDJ: IS
THIS TRUE?  DWH, Currently I'm not using a jitter term, but I can and will) in what follows.
In this formulation, if $\theta$ are the parameters of interest, the
jitter variance $s^2$ is a nuisance noise-model parameter.
The Gaussian Process will be a generalization of this modification.

@@

We are now left to demonstrate that it is possible to infer the
parameters of interest $\theta$ while marginalizing over the
calibration parameters $\alpha$.  We will apply the methodology for
spectrophotometric calibration described above to an \emph{actual,
real} optical spectrum of a stellar cluster in M31.  Spectra of star
clusters provide a good test case, as they are less complex than
galaxy spectra (for which it is necessary to model the star formation
history) but are by no means trivial. In this way the key features of
the calibration model and methodology are not obscured.

We model the stellar cluster as a single age SSP. The spectrum of a
star cluster under this assumption is given by
\begin{eqnarray}\label{eq:MeanModel} 
\mu(\theta, \lambda(1+z)) & = &
\frac{m_*}{4\pi d_L^2(1+z)} G(\lambda,\sigma_v) \otimes 
S(\lambda, t, Z [, \beta]) e^ {-k_\lambda\tau_V} 
\\ 
\theta & \equiv & [m_*, t, Z, \tau_V, \sigma_v, z, d_L, \beta]
\quad ,
\end{eqnarray} where the parameters $\theta$ include include the age
$t$, metallicity $Z$, stellar mass $m_*$, dust attenuation effective
$V$-band optical depth $\tau_V$, velocity dispersion $\sigma_v$,
luminosity distance $d_L$ and redshift $z$.  The spectrum of one solar
mass of stars that formed at time $t$ with metallicity $Z$ is given by
$S(\lambda, t, Z [, \beta])$ where the additional parameters $\beta$
may include details of the population synthesis, such as the IMF slope
and adjustments to the stellar spectra or stellar evolutionary tracks
\citep[e.g.][]{FSPS}. We will assume $\beta$ to be fixed for the
purposes of this exercise, with no loss of generality (DO WE WANT TO
VARY IMF SLOPE?).  We will also fix the shape of the attenuation
curve, $k_\lambda \equiv \tau_\lambda/\tau_V$ and assume that $d_L$ is
known to high accuracy from independent methods.

To model the photometry we simply project the observed frame mean
model spectrum $\mu(\theta, \lambda(1+z))$ onto the filters of
interest.  Note that we do \emph{not} include the calibration vector
$f(\alpha)$ in this projection.

BDJ/DRW WORDS ABOUT DATA

To demonstrate the effectiveness of our spectrophotometric calibration
strategy we will infer the parameters $\theta$ from the spectrum both
\emph{before} and \emph{after} it has been spectrophotometrically
calibrated using typical techniques and ancillary spectrophotometric
calibration data. (PROBABLY NEED WORDS FROM NELSON HERE, OR CITE HIS
PAPER IF DETAILED ENOUGH)

@@

BDJ INSERT WORDS HERE.
BDJ random issues, collected here in case we want to address them and
to keep me from cluttering up the main text: effective aperture
corrections (i.e. we're implicitly assuming photometry and
spectroscopy sample the same population).  Sky subtraction residuals.
Model resolution and instrument resolution.  Method only useful when
you have a model for the mean spectrum?  How to interpret GP
residuals, as a fn of S/N?  Photometric upper limits?
